---
editor_options:
  markdown:
    wrap: 72
output: pdf_document
---

## Chapter 11
**(6)** Prove parts (b) and (c) 
(b) $\widehat{T}_r c_k=u_n^{-r k} \hat{c}_k$.

\begin{align*}
\widehat{T}_r c_k &= \frac{1}{n}\Sigma^{n-1}_0T_rc_je^{-2 \pi ijk / n}\\
&\overset{(*)}{=}\frac{1}{n}\Sigma^{n-1}_0 c_{j-r} e^{-2 \pi ijk / n}\\
&= \frac{1}{n}\Sigma^{n-1}_0 c_{j-r}e^{-2 \pi i(j-r)k / n} e^{-2 \pi irk / n}\\
&\overset{(**)}{=}u_n^{-r k} \hat{c}_k
\end{align*}

(*) holds because $T_r c_j=c_{j-r}$ 
(**) holds because $\hat{c}_{j-r}=\frac{1}{n} \sum_{j=0}^{n-1} c_{j-r} e^{-2 \pi i \frac{(j-r) k}{n}}$ and $u_n=e^{2 \pi i / n}$

(c) $\widehat{R c}_k=R \hat{c}_k$.
\begin{align*}
\widehat{R c}_k&=\frac{1}{n} \sum_{j=0}^{n-1}R c_j e^{-2 \pi i \frac{j k}{n}}\\
&\overset{(*)}{=}\frac{1}{n} \sum_{j=0}^{n-1}c_{-j}e^{-2 \pi i \frac{j k}{n}}\\
& = R\hat{c}_k
\end{align*}
(*) holds because $R c_j=c_{-j}$ 

**(19)** The $n \times n$ permutation matrix $\boldsymbol{S}_n$
satisfying $\boldsymbol{S}_n \boldsymbol{e}_k=\boldsymbol{e}_{k+1}$ for
$k=1, \ldots, n-1$ and
$\boldsymbol{S}_n \boldsymbol{e}_n=\boldsymbol{e}_1$ is called a
circular shift. Prove that the $n \times n$ matrix $C$ is circulant if
and only if $S_n C=C S_n$.\
$\textit{Solution}:$ the $ij$th entry of circular shift S is given by
$S_{i j}=1(i=j+1)$ In particular, the left (right) multiplication by S
equals to row (column) circular permutation,\
so for any circulant matrix $C$ we can have $SC = CS$ note we have  
\begin{align*}
& (S C)_{i j}=\sum_{\ell} S_{i \ell} C_{\ell j}=C_{i-1, j} \\
& (W C)_{i j}=\sum_{\ell}^{\ell} C_{i \ell} S_{\ell i}=C_{i, j+1}
\end{align*}

Then we get\
$$
{i-1, j}=C_{i, j+1} \Leftrightarrow C_{i, j}=C_{i-1, j-1} \Leftrightarrow C_{i j}=C_{i+k, j+k} 
$$
Hence the matrix $C$ is a circulant matrix\
Proved

**(20)** Prove that the sum or product of two circulant matrices of the
same size is circulant. Also prove that the inverse of an invertible
circulant matrix C is circulant. Thus, the collection of invertible
circulant matrices forms a group. Why is this group commutative? 
$\textit{Solution}:$ A matrix is circulant if and only if it commutes
with 
$$
J=\left(\begin{array}{ccccc}
0 & 1 & 0 & \cdots & 0 \\
0 & 0 & 1 & \cdots & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots \\
0 & 0 & 0 & \cdots & 1 \\
1 & 0 & 0 & \cdots & 0
\end{array}\right)
$$ 
So the sum or product of two circulant matrices of the same size is circulant too. 
By Cayley-Hamilton theorem, the inverse of every nonsingular matrix $A$ can be expressed as a polynomial in $A$ of degree $\leq n-1$.
Therefore, if $A$ is in the linear span of $I, J, \ldots, J^{n-1}$, then
$A^{-1}$ is a polynomial in $J$ of degree $\leq(n-1)^2$.\
Since $J^n=I$, it follows that $A^{-1}$ lies inside the linear span of
$I, J, \ldots, J^{n-1}$.  
Hence it is circulant.
Circulant matrices commute because they are simultaneously diagonaliz-able.


## Chapter 12

```{r}
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(grid)
library(MASS)
```

**(1)** Discuss how you would use the inverse method to generate a
random variable with (a) the continuous logistic density 
$$
f(x \mid \mu, \sigma)=\frac{e^{-\frac{x-\mu}{\sigma}}}{\sigma\left[1+e^{-\frac{x-\mu}{\sigma}}\right]^2}
$$ 
(b) the Pareto density 
$$
f(x \mid \alpha, \beta)=\frac{\beta \alpha^\beta}{x^{\beta+1}} 1_{(\alpha, \infty)}(x),
$$ 
and (c) the Weibull density 
$$
f(x \mid \delta, \gamma)=\frac{\gamma}{\delta} x^{\gamma-1} e^{-\frac{x^\gamma}{\delta}} 1_{(0, \infty)}(x),
$$
where $\alpha, \beta, \gamma, \delta$, and $\sigma$ are taken positive.\
$\textit{Solution}:$\
Step 1: Generate $u$ from uniform $(0,1)$\
Step 2: Return $x=F_X^{-1}(u)$

```{r}
# generate 10,000 random uniform variables
set.seed(1)
u <- runif(10000)
```

(a) continuous logistic density

$$ F_X(x) = \frac{1}{1+e^{-\frac{x-\mu}{\sigma}}} = u $$
$$\frac{1}{u} = 1+e^{-\frac{x-\mu}{\sigma}}$$
$$\frac{1-u}{u} = e^{-\frac{x-\mu}{\sigma}}$$
$$\log(\frac{1-\mu}{\mu})=\frac{\mu -x}{\sigma}$$ 
$$F_X^{-1} = x = \mu - \log(\frac{1-u}{u}) \cdot \sigma $$ 
Take an example with $\mu=0$ and $\sigma = 1$

```{r}
mu = 0
sigma = 1
generated_vals <- floor( mu - log( (1-u)/u )*sigma )
plts <- list()


x_dlogis <- seq(- 10, 10, by = 0.1) 
y_dlogis <- dlogis(x_dlogis,location = 0, scale = 1)

# construct histogram of generated values
plts[[1]] <- tibble(generated_vals = generated_vals) %>%
  ggplot(aes(x = generated_vals)) +
        geom_histogram(binwidth = 0.5 ) + 
        theme_bw() +
        labs(x = "x", y = "Frequency", title = "Generated continuous logistic density")

plts[[2]] <- tibble(x = x_dlogis, theoretical_pdf = y_dlogis) %>%
    ggplot() +
        geom_line(aes(x = x, y = theoretical_pdf), colour = "blue") +
        theme_bw() +
        labs(x = "x", y = "f(x)", title = "Exact continuous logistic density")

grid.arrange(grobs = plts, nrow = 1)
```

(b) Pareto density
$$F_X(x)=1-\left(\frac{\alpha}{x}\right)^\beta ; x \geq \alpha$$

$$
F(X)=U 
$$

$$U \sim\text { Uniform }(0,1)$$
$$1-\left(\frac{\alpha}{x}\right)^\beta=u $$
$$F_X^{-1} = x=\alpha(1-u)^{-1 / \beta}$$

(c) Weibull density\
\begin{align*}
F_X(x) &=u=1-\exp \left(\frac{x^\gamma}{\delta}\right), \text { where } \mathrm{x} \geq 0  \\
& \Longrightarrow \log (1-u)=\frac{x^\gamma}{\delta} \\
& \Longrightarrow \log (1-u) \delta=x^\gamma . \\
F_X^{-1} &=x=\left(\log (1-u) \delta\right)^{1 / \gamma} \\
&=\log (1-u)^{1 / \gamma} \cdot\left(\delta\right)^{1 / \gamma}
\end{align*}

**(2)** Continuing problem 1, discuss how the inverse method applies to
(d) the Gumbel density 
$$
f(x)=e^{-x} e^{-e^{-x}},
$$
(e) the arcsine density 
$$
f(x)=\frac{1}{\pi \sqrt{x(1-x)}} 1_{(0,1)}(x),
$$
and (f) the slash density 
$$
f(x)=\alpha x^{\alpha-1} 1_{(0,1)}(x),
$$
where $\alpha>0$.

$\textit{Solution}:$\
Step 1: Generate $u$ from uniform $(0,1)$;\
Step 2: Return $x=F_X^{-1}(u)$\
(d) Gumbel density\
$$F_X(x) = e^{-e^{(-x)}}=u$$ $$-e^{(-x)}=\log(u)$$
$$-x =\log(-\log(u))$$ $$F_X^{-1} = x = -\log(-\log(u))$$ (e) arcsine
density

$$F_X(x)=\frac{2}{\pi} \arcsin (\sqrt{x}) = u$$
$$\arcsin (\sqrt{x}) = u\cdot \frac{\pi}{2}$$
$$F_X^{-1} = x  = \sin^2(u\cdot \frac{\pi}{2})$$

(f) slash density The slash density I found on the Internet is different
    from here.\
    $$F_X(x)=\int_0^x f(t)dt = \int_0^x \alpha t^{\alpha-1} d t=x^\alpha =u$$
    $$F_X^{-1} = x = u^{1/\alpha}$$

**(5)** Describe one method for generating independent Poisson deviates
and implement it in code.\
$\textit{Solution}:$\
$Y \sim \operatorname{Pois}(\lambda)$\
1. Define parameter $\lambda$ =5.\
2. Generate $k$ random uniform variables $U \sim U n i f[0,1]$.\
3. Derive an expression for $y$.\
The probability mass function is
$P\{Y=y\}=\frac{e^{-\lambda} \lambda^y}{y !}$ for $y=0,1,2, \ldots$\
We can recursively define this with $p_0=e^{-\lambda}$ and
$p_{i+1}=\frac{\lambda}{i+1} p_i$.\
The cumulative distributive function can be expressed as
$F(y)=\sum_{i=0}^y p_i$.\
We want $y$ to be the value of $j$ which
$\sum_{i=0}^{j-1} p_i \leq u<\sum_{i=0}^j p_i$.\
4. We define a function that uses the recursive expression of $p_i$ to
build the CDF until we obtain $y$.\
For the base case $i=0$, we define $p_0$ and initiate $F(y)$. We
increment $i$ and update $F(y)$ with the new $p_i$ until we find $y$.

```{r}
#Define parameter
lambda =5
#Generate random uniform variables
u <- runif(10000)
#set function
poisgen = function(u){
    y = 0
    i = 0
    p = exp(-lambda)
    F = p
    while(u>=F){
        p = (p*lambda)/(i+1)
        F = F+p
        i = i+1
        y = i
    }
    return(y)
}
ysim <- sapply(u, poisgen)

pois <- dpois(1:15, lambda)

# create list to hold our two plots
plts <- list()

plts[[1]] <- tibble(sim=ysim) %>% ggplot() +
  geom_bar(aes(x=as.factor(sim)), stat="count") +
  labs(x="y", y="Frequency", title="Simulated Poisson PMF") +theme_bw()

plts[[2]] <- tibble(x=1:15, pmf=pois) %>% ggplot() +
  geom_bar(aes(x=as.factor(x), y=pmf), stat="identity") +
  labs(x="y", y="P(Y=y)", title="Exact Poisson PMF") +theme_bw()

grid.arrange(grobs = plts, nrow = 1)

```

**(6)** Describe and implement acceptance-rejection sampling for the
beta distribution with parameters $\alpha>1$ and $\beta>1$. In this case
$x^{\alpha-1}(1-x)^{\beta-1} \leq c$ for all $x \in[0,1]$ and some
constant $c$. First, calculate the least value of $c$.\
Generate $Y \sim \operatorname{Beta}(\alpha=2.7, \beta=6.3)$\
1. Generate $(U, V)$ independent uniform $(0,1)$.\
2. If $U<\frac{1}{c} f_Y(V)$ then, set $Y=V$; otherwise, return to
step 1. Where $c=\max _y f_Y(y)=2.669$

```{r}
f1 = function(x){
  a = 2.7
  b = 6.3
  beta = gamma(a) * gamma(b) /gamma(a + b)
  p = x ** (a - 1) * (1-x) ** (b - 1)
  return (1/beta * p)
}
mode = (2.7-1)/(2.7+6.3-2) ##Mode of Beta distribution
c = f1(mode)
print(c)
```

```{r}
beta_gen = function(n){
  set.seed(199)
  i = 0
  output = c()
  while (i < n){
    U = runif(1)
    V = runif(1)
    if( U < 1/c * f1(V)){
      output[i] = V
      i = i + 1
  }
  }
  return (output)
}

betagen = beta_gen(n = 100000)

plts <- list()

plts[[1]] <- tibble(sim = betagen) %>%
  ggplot() + geom_density(aes(x = sim)) + labs(x = "x", y = "Frequency", title =
                                                 "Generated Beta density") +
  xlim(0,1)  + theme_bw()

gam <- dbeta(seq(0, 1,0.01), shape1 = 2.7,shape2 = 6.3,ncp = 0)
# construct line chart of exact pdf
plts[[2]] <- tibble(x = seq(0, 1,0.01), pdf = gam) %>%
  ggplot() + geom_line(aes(x = x, y = pdf)) + labs(x = "x", y = "f(x)", title =
                                                     "Exact Beta density") + 
  xlim(0, 1)  + theme_bw()

#arrange two plots
grid.arrange(grobs = plts, nrow = 1)
```
